---
title: "RHEAS preliminary maize yield prediction visualization"
author: "Benson Kenduiywo"
date: "08/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

Load data

```{r d1, echo=FALSE}
#df <- read.csv('D:/RCMRD/Code/Rheas_Yields/kenya_1990_2022.csv')
#head(df,n =5)
df <- read.csv('D:/RCMRD/Code/Rheas_Yields/uganda_1990_2022.csv')
```


## Data preparation

Extract Harvest & plant dates into months and years

```{r d2, echo=FALSE}
df$harvest <- as.Date(df$harvest)
df$planting <- as.Date(df$planting)
df$h_month  <- as.numeric(format(df$harvest,"%m"))
df$h_year  <- as.numeric(format(df$harvest,"%Y"))
df$p_month  <- as.numeric(format(df$planting,"%m"))
df$p_year  <- as.numeric(format(df$planting,"%Y"))
head(df, 5)
```

Let's define/extract the various seasons in Kenya.

```{r s1, echo=FALSE}
#Long Rain Long Dry (LRLD) season
#df$season <- df[df$p_month > 1 & df$p_month < 6, ]
df$season <- NA
df$season[df$p_month > 1 & df$p_month < 6] <- 'LRLD'
#Short Rain Short Dry (SRSD)
df$season[df$p_month > 9 & df$p_month <= 12] <- 'SRSD'
#sd <- df[df$p_month > 9 & df$p_month <= 12,]

```


lets now aggregate RHEAS estimated yields per year per county.

```{r y1, echo=FALSE}
#lr <- aggregate(df[,'gwad'], df[ , c("cname", "h_year")], mean, na.rm=TRUE)
yields <- aggregate(gwad~cname+h_year, df, mean, na.rm=TRUE)
names(yields)[1] <- 'District'
write.csv(yields, 'D:/RCMRD/Code/Rheas_Yields/uganda_aggregated_1990_2022.csv')

```


## Plots

Annual maize yields variation.

```{r y2, echo=FALSE}
x11()
boxplot(gwad~h_year, data=yields, xlab='Year', ylab = 'Maize yields (kg/ha)')
```

Country level yields.

```{r y3, echo=FALSE}
x11()
c <- aggregate(gwad~h_year, df, mean, na.rm=TRUE)
c$gwad <- c$gwad/1000
barplot(gwad~h_year, data=c, xlab='Year', ylab = 'Maize yields (MT/ha)', main="Country's total production", ylim = c(0,2))

```


## Display maps

Load country shapefile

```{r ds1, echo=FALSE}
library(raster)
ug <- shapefile('uga_admbnda_adm2_ubos_20200824.shp') #Districts
#ug <- getData("GADM",country="UGA",level=2) #Boundaries are old
names(ug)[3] <- 'District'
proj <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
crs(ug) <- proj
library(tmap)
tmap_mode("view")
tm_shape(ug, name="Region") +
  tm_polygons('ADM1_EN', title='Region', id='ADM1_EN', legend.show = TRUE)+
  tm_shape(ug, name="District") +
  tm_borders(col = 'black',lwd = 0.75)+
  tm_text("District", size = 0.9)

```

Merge boundary with average distict level yield information.

```{r ds2, echo=FALSE}
d <- aggregate(gwad~District, yields, mean, na.rm=TRUE)
dff <- merge(ug[,"District"],d, by="District")
```

Visualize average district level information.

```{r ds3, echo=FALSE}
tmap_mode("view")
tm_shape(dff, name="Average yield (kg/ha) per District") +
  tm_polygons("gwad", id="gwad",breaks = c(100, 500, 1000, 1500, 2000,2500),  title="Maize yield (kg/ha) 1990-2021") +
  tm_text("District", size = 0.9)+
  tm_format("World")

```
